Analyzing Influencing Factors on Family Size in Portugal

### 2025-02-10 data/data_cleaning.Rmd
Update

for(D in intersect(names(portugal), names(formats))){
  		portugal[[D]] = factor(portugal[[D]],
  			levels=formats[[D]]$code, 
				labels=formats[[D]]$label)
}
portugal$ageMarried = relevel(portugal$ageMarried, '22to25')
portugal$region = relevel(portugal$region, 'lt10k')

if(TRUE) save(portugal, file='raw_data.RData')
```

```{r thedata}
head(portugal)
table(portugal$region)
```

>>>>

for(D in intersect(names(portugal), names(formats))){
  		portugal[[D]] = factor(portugal[[D]],
  			levels=formats[[D]]$code, 
				labels=formats[[D]]$label)
}
portugal$region = relevel(portugal$region, 'lt10k')

if(TRUE) save(portugal, file='Portugal.RData')
```

```{r thedata}
head(portugal)
table(portugal$ageMarried)
```

### 2025-02-10 paper/paper.rmd
Update

library(tidyverse)
```

@@

To analyze this relationship, we will use Generalized Linear Models (GLMs) as they are well-suited for count data like family size. Specifically, Poisson regression or negative binomial regression will be considered based on the presence of overdispersion. Exploratory data analysis will be conducted to summarize key variables, followed by model selection techniques to identify the best-fitting statistical model. R [@r] will be used to conduct this, and packages including `tidyverse` [@tidyverse], `here` [@here], `knitr` [@knitr], and `MASS` [@MASS] will also be used. The final model will be interpreted and draw meaningful conclusions about the relationship between literacy, age at marriage, and family size in Portugal.

@@

>>>>

library(tidyverse)
library(here)
library(knitr)
library(gridExtra)
library(patchwork)
library(MASS)
library(glmmTMB)
load(here::here("data/Portugal.RData"))
```

@@

To analyze this relationship, we will use Generalized Linear Models (GLMs) as they are well-suited for count data like family size. Specifically, Poisson regression or negative binomial regression will be considered based on the presence of overdispersion. Exploratory data analysis will be conducted to summarize key variables, followed by model selection techniques to identify the best-fitting statistical model. R [@r] will be used to conduct this, and packages including `tidyverse` [@tidyverse], `here` [@here], `knitr` [@knitr], `kableExtra` [@kableExtra], `patchwork` [@patchwork], `xMASS` [@MASS], and `glmmTMB` [@glmmTMB] will also be used. The final model will be interpreted and draw meaningful conclusions about the relationship between literacy, age at marriage, and family size in Portugal.

@@

```{r}
#| label: tbl-summarize
#| tbl-cap: "1"
#| echo: false

kable(portugal |>
        summarize(Mean = round(mean(children), 1),
                  SD = round(sd(children), 1),
                  Median = median(children),
                  Minimum = min(children),
                  Maximum = max(children),
                  IQR = round(IQR(children), 1)))

kable(portugal |>
        count(literacy) |>
        rename("Literacy Status" = literacy, "Count" = n))

kable(portugal |>
        count(ageMarried) |>
        rename("Age at Marriage" = ageMarried, "Count" = n))

grid.arrange(table1, table2, table3, ncol = 2)
```

```{r, fig.pos="H", fig.height=6, fig.width=8}
#| label: fig-bar
#| fig-cap: "Overview of key demographic variables: number of children in a family, literacy status, and age at marriage. The left panel (blue) displays the distribution of the number of children per family, showing a right-skewed pattern where most families have fewer children, with the highest frequency at 1 to 3 children. The top-right panel (green) represents literacy status, categorized as 'yes' (literate) or 'no' (illiterate). The majority of individuals are literate, with a much smaller proportion classified as illiterate. The bottom-right panel (red) illustrates the distribution of age at marriage, indicating that the most common age range for marriage is 22 to 25 years, followed by 20 to 22 and 25 to 30 years. Fewer individuals marry at very young (0 to 15, 15 to 18) or older ages (30 and above). These distributions provide critical context for subsequent Generalized Linear Model (GLM) analysis, as they highlight demographic patterns that may influence relationships between these variables."
#| echo: false

bar1 <- ggplot(portugal, aes(x = children)) +
  geom_bar(fill = "blue", alpha = 0.7) +
  theme_minimal() +
  labs(x = "Age at Marriage (Years)", y = "Number of Individuals")

bar2 <- ggplot(portugal, aes(x = literacy)) +
  geom_bar(fill = "green", alpha = 0.7) +
  theme_minimal() +
  labs(x = "Age at Marriage (Years)", y = "Number of Individuals")

bar3 <- ggplot(portugal, aes(x = ageMarried)) +
  geom_bar(fill = "red", alpha = 0.7) +
  theme_minimal() +
  labs(x = "Age at Marriage (Years)", y = "Number of Individuals")

bar1 + bar2 /bar3
```
